---
title: "Testing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteIndexEntry{Testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

```{r setup, include = FALSE}
library(inedata)
library(dplyr)
```

# Package Tutorial

The `inedata` package allows to download datasets produced by INE Chile. The package communicates with an API REST, which makes it possible to retrieve many household survey datasets using R code. All the available data provided by this API corresponds to files published on INE's web site. 

At the moment the API is restricted to the institutional network. Because of that the package works only in this environment. We hope to make the API available to the general public as soon as possible.     

## First steps

At the moment the package can be installed from github or gitlab using devtools.  

```{r, eval=FALSE}
# From github
devtools::install_github("inesscc/dataine")

# From gitlab
devtools::install_gitlab("inesscc/dataine")
```

## Getting catalog

`get_catalog()` returns a dataframe containing all the datasets available in the repository and its versions when the `dataset` parameter is not specified. The following datasets are currently available:  

* *EPF*
* *ENUSC*
* *ESI*
* *ENE*

It also adds summary data in the form of number of columns and rows per dataset (in the *n_col* and *n_row* columns, respectively). The *survey* column contains the survey where the dataset comes from. The *version* column indicates the year or quarter when the dataset was collected.       

```{r}
get_catalog()
```
The `dataset` parameter retrieves information for a specific survey. In this example we are using *ene*, which returns all the available versions for that survey. We will store the file version for the june-august-september quarter of 2011 (2011-08-jas) in a variable called `file`.

```{r}
get_catalog(dataset = "ene")

file <-  get_catalog(dataset = "ene") %>% 
  filter(version=="2011-08-jas") %>% 
  pull(version)
```

## Getting columns

The `get_columns()` function returns the columns for a specific file. It requires two mandatory parameters, namely `dataset` and `version` to fetch the specific column names.
We already have the name of one of the available versions of the *ene* survey. 

```{r}
get_columns(dataset = "ene", version = file)
```


## Downloading data

The `get_data()` function allows us to bring a dataset from a specific survey. The function has four parameters: `dataset`, `version`, `col_list` and `save_where`.
For instance, we can get the *ene* data for a specific quarter. If the `version` parameter is not specified then it will return the newest version of the survey. 

```{r, message=FALSE,warning=FALSE,results='hide'}
# Get dataset of a specific version
ene <-  get_data(dataset = "ene", version = "2010-02-efm" ) 

# Get newest version
ene2 <- get_data(dataset = "ene") 
```
```{r, eval=TRUE}
ene %>% head(3)
```

With `col_list` we can select some of the variables in the dataset, if not specified it will select all columns. Meanwhile, `save_where` allows us to define where the dataset will be stored. There are four options:  

* **renviron**: store in the memory of R enviroment. If the size overloads the ram memory it will show a warning and a prompt recommending to save the data in the working directory and asking if the user wants to proceed.  
* **disk**: store in a directory called *data/* in the working directory, but not in the R enviroment.
* **both**: store in both parts.
* **no_message**: under the users responsibility, it will not show the warning and the datasets will be stored in R enviroment.

```{r, message=FALSE,warning=FALSE,results='hide'}
ene_small <- get_data(dataset = "ene", version = file, col_list = c("ano_trimestre","mes_central","fact_cal","cae_especifico"), save_where = "no")
```
```{r}
ene_small %>% head()
```

**Important note:** we highly recommend against downloading the entire dataset. Instead, consider selecting specific columns, as the full download could consume a substantial amount of time and storage space. 

The `get_many_data()` function allows to download more than one dataset at the same time. In this example we are getting the versions between december-january-february (2022-01-def) and february-march-april (2022-03-fma) 2022 quarters of the *ene* survey.   
The function has six parameters: `dataset`, `from`, `to`, `col_list`, `save_where`, `memory_warning_limit`.
The `from`and `to` arguments are mandatory and are used to filter by the selected periods or versions of a particular survey.
The `memory_warning_limit` argument allows you to set the memory limit that triggers the warning when downloading to R memory.  
The output is a list of *dataframes*. In this case we can see the *df_list* contains 3 elements.

```{r, message=FALSE,warning=FALSE,results='hide'}
df_list <-  get_many_data(dataset = "ene", from = "2022-01-def", to = "2022-03-fma" )
```
```{r}
length(df_list)
```

We can convert the three *dataframes* into one table

```{r}
ene_table <-  df_list %>%
  bind_rows()
```
**Important note:** different versions of the same dataset are not always compatible, so, the `bind_rows()` method won't always work. Be always careful when joining multiple versions and check their structure before proceeding.

# API Tutorial

The R package works with a service developed in Python, whose documentation can be found
[here](http://10.90.10.46:7000/docs).Below are the endpoints and a link to the documentation.

**Important note:** when interacting directly with the API, it is important to be aware that due to data format restrictions, any cells that were originally NA will be displayed as empty text.


## Returns available datasets

A *.json* file is entered containing the survey, version and optionally a list of columns. If a list of columns is not entered, the API returns a table with all the columns of the survey. The final output is also in a *.json* format.  
The next example shows how to get the dataset of the *enusc* 2017 survey selecting the columns "P1_1_1" and "P3_1_1" in a *.json* format.

```{r, eval=FALSE}
{
  "dataset": "enusc",
  "version": "2017",
  "col_list": [
    "P1_1_1",
    "P3_1_1"
  ]
}
````

The endpoint is the following:

```{r, eval=FALSE}
http://10.90.10.46:7000/data
```

## Returns the columns of a file

The user has to enter the dataset and version, then the API will return a vector containing the columns of the selected dataset.

The endpoint is the following:

```{r, eval=FALSE}
http://10.90.10.46:7000/colnames
```

## Returns summary data

Delivers metadata of the available tables. Optionally, you can enter a survey to filter using the parameter *dataset_filter*, obtaining the metadata of the tables associated with that survey in *.json* format. Data includes dataset name, dataset version, number of rows, number of columns and file size in bytes.

The endpoint is the following:

```{r, eval=FALSE}
http://10.90.10.46:7000/info
```

